---
title: "Support"
author: "Kimberly A. Brink"
date: '`r format(Sys.Date(),"%B %d, %Y")`'
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

```{r install_libraries}
library(lubridate)
library(Hmisc)
```

```{r standard_error}
s.error <- function(x) sd(x)/sqrt(length(x))
```

```{r load_file}
#detach(support)
#filename="/Volumes/lsa-research01/ALL STUDIES/Current Studies/Support/Support 3.0/Support - Data v3.0.sav"
#support <- spss.get(filename,datevars=c("DOB","DOT"))

filename="/Volumes/lsa-research01/ALL STUDIES/Current Studies/Support/Support 3.0/Support - Data v3.0.sav"
support <- spss.get(filename,datevars=c("DOB","DOT"))

```

```{r calculate_variables}
support$Age = (year(as.period(new_interval(support$DOB, support$DOT)))*12) + month(as.period(new_interval(support$DOB, support$DOT))) + (day(as.period(new_interval(support$DOB, support$DOT)))/30)
support = support[which(support$DOT>='2016-01-13'),]

support$Crawl2 = ifelse(support$Crawl==levels(support$Crawl)[5], 3, ifelse(support$Crawl==levels(support$Crawl)[4], 2, ifelse(support$Crawl==levels(support$Crawl)[3], 1, ifelse(support$Crawl==levels(support$Crawl)[2], 0, NA))))
```

###Demographics.
There were `r length(support$SubID)` participants.

```{r computer_error}
support$ContactTest.Average = (support$ContactTest1+support$ContactTest2+support$ContactTest3)/3
support$NoContactTest.Average = (support$NoContactTest1+support$NoContactTest2+support$NoContactTest3)/3

support = support[which(support$ContactTest.Average!='NA'),]
support = support[which(support$NoContactTest.Average!='NA'),]
```

```{r computer_error2}
#support = support[which((support$ContactTest1 >= 4)&(support$ContactTest2 >= 4)&(support$ContactTest3 >= 4)&(support$NoContactTest1 >= 4)&(support$NoContactTest2 >= 4)&(support$NoContactTest3 >= 4)),]

support = support[which((support$ContactTest1 >= 4)&(support$NoContactTest1 >= 4)),]
#support = support[which((support$ContactTest2 >= 4)&(support$NoContactTest2 >= 4)),]
#support = support[which((support$ContactTest3 >= 4)&(support$NoContactTest3 >= 4)),]
```

```{r outliers}
# averages = append(support$ContactTest.Average, support$NoContactTest.Average)
# support = support[which(support$ContactTest.Average <= (mean(averages)+(2*sd(averages)))),]
# support = support[which(support$NoContactTest.Average <= (mean(averages)+(2*sd(averages)))),]
# support = support[which(support$ContactTest.Average >= (mean(averages)-(2*sd(averages)))),]
# support = support[which(support$NoContactTest.Average >= (mean(averages)-(2*sd(averages)))),]
```

Now there are `r length(support$SubID)` participants after removing for computer error and outliers. 

```{r crawling}
#support = support[which(support$Crawl2>0),]
#support = support[which(support$Age>=6.5),]
```

Now there are `r length(support$SubID)` participants after removing for crawling. 

```{r demographics}
attach(support)

isGirl <- support$Sex == "Female"
```

The average age of the sample is `r round(mean(Age), digits = 2)` months.

The minimum age of the sample is `r round(min(Age), digits = 2)` months.

The maximum age of the sample is `r round(max(Age), digits = 2)` months.

There are `r sum(isGirl)` females in the sample.

The first date of test was `r min(DOT)`.

The most recent date of test was `r max(DOT)`.

#Tests for normality
```{r boxplots}
#plot(c(NoContactTest1, NoContactTest2,NoContactTest3,ContactTest1,ContactTest2,ContactTest3))
```
```{r histograms, fig.width = 8, fig.height = 5}
hist(NoContactTest1)
hist(ContactTest1)
hist(NoContactTest2)
hist(ContactTest2)
hist(NoContactTest3)
hist(ContactTest3)
```

#Bar Graph Within, All participants
```{r barplot, fig.path='Figures/', dev=c('png','postscript'), fig.width = 6, fig.height = 6}
medians <- matrix(c(median(ContactTest1),median(ContactTest2),median(ContactTest3),median(NoContactTest1),median(NoContactTest2),median(NoContactTest3)),ncol=3,byrow=TRUE)
colnames(medians) <- c("Test Trial 1","Test Trial 2","Test Trial 3")
rownames(medians) <- c("Contact","No Contact")
medians <- as.table(medians)

g.range = range(0,ceiling(max(medians))+5)

plot.new()
barCenters <- barplot(medians, col=c("darkblue","red"), legend = rownames(medians), ylab = "Median Looking Time (sec)", beside = TRUE, ylim=g.range)
text(barCenters,medians+1.5,label=medians)
```

#Wilcox Test
Because the distribution of the Test trial is not normal we will use the non-parametric within samples t-test.
```{r}

t1 = wilcox.test(NoContactTest1,ContactTest1,paired=TRUE)
Z1 = round(qnorm(t1$p.value/2),digits = 2)
p1 = round(t1$p.value, digits = 2)
p1 <- ifelse(p1<0.001, 0.001, ifelse(p1<0.01, 0.01, ifelse(p1<0.05,0.05,p1)))

 t2 = wilcox.test(NoContactTest2,ContactTest2,paired=TRUE)
 Z2 = round(qnorm(t2$p.value/2),digits = 2)
 p2 = round(t2$p.value, digits = 2)
 p2 <- ifelse(p2<0.001, 0.001, ifelse(p2<0.01, 0.01, ifelse(p2<0.05,0.05,p2)))

 t3 = wilcox.test(NoContactTest3,ContactTest3,paired=TRUE)
 Z3 = round(qnorm(t3$p.value/2),digits = 2)
 p3 = round(t3$p.value, digits = 2)
 p3 <- ifelse(p3<0.001, 0.001, ifelse(p3<0.01, 0.01, ifelse(p3<0.05,0.05,p3)))
```

There was no difference in looking times between the first Contact test event (Median =
`r median(ContactTest1)`) and the first No Contact event (Median = 
`r median(NoContactTest1)`),  (Wilcoxon *Z* = `r Z1`,  *p* = `r p1`)

There was no difference in looking times between the second Contact test event (Median = `r median(ContactTest2)`) and the second No Contact event (Median = 
`r median(NoContactTest2)`),  (Wilcoxon *Z* = `r Z2`,  *p* = `r p2`).

There was a difference in looking times between the third Contact test event (Median =
`r median(ContactTest3)`) and the third No Contact event (Median = 
`r median(NoContactTest3)`),  (Wilcoxon *Z* = `r Z3`,  *p* = `r p3`).

#Bar Graph Means
```{r barplot_means_65, fig.path='Figures/', dev=c('png','postscript'), fig.width = 8, fig.height = 7}
means <- matrix(c(mean(ContactTest1),mean(ContactTest2),mean(ContactTest3),mean(NoContactTest1),mean(NoContactTest2),mean(NoContactTest3)),ncol=3,byrow=TRUE)
se <- matrix(c(s.error(ContactTest1),s.error(ContactTest2),s.error(ContactTest3),s.error(NoContactTest1),s.error(NoContactTest2),s.error(NoContactTest3)),ncol=3,byrow=TRUE)

colnames(means) <- c("Test Trial 1","Test Trial 2","Test Trial 3")
rownames(means) <- c("Contact","No Contact")
means <- as.table(means)
meansR <- round(means,2)
g.range = range(0,ceiling(max(means))+10)

plot.new()
barCenters <- barplot(means, col=c("darkblue","red"), legend = rownames(means), ylab = "Mean Looking Time (sec)", beside = TRUE, ylim=g.range)
arrows(barCenters, means-se, barCenters, means+se, lwd = 2, length=0.05, angle=90, code=3)
text(barCenters,means+se+1.5,label=meansR)
```

```{r ttest_65}
t1 = t.test(NoContactTest1,ContactTest1,paired=TRUE)

t2 = t.test(NoContactTest2,ContactTest2,paired=TRUE)

t3 = t.test(NoContactTest3,ContactTest3,paired=TRUE)
```

There was no difference in looking times between the first Possible test event (M =
`r round(mean(ContactTest1), 2)`) and the first Impossible event (M = 
`r round(mean(NoContactTest1),2)`),  (t(`r round(t1$parameter,2)`) = `r round(t1$statistic,2)`,  *p* = `r round(t1$p.value,2)`).

There was no difference in looking times between the second Possible test event (M = 
`r round(mean(ContactTest2),2)`) and the second Impossible event (M = 
`r round(mean(NoContactTest2),2)`),  (t(`r round(t2$parameter,2)`) = `r round(t2$statistic,2)`,  *p* = `r round(t2$p.value,2)`).

There was a difference in looking times between the third Possible test event (M =
`r round(mean(ContactTest3),2)`) and the third Impossible event (M = 
`r round(mean(NoContactTest3),2)`),  (t(`r round(t3$parameter,2)`) = `r round(t3$statistic,2)`,  *p* = `r round(t3$p.value,2)`).

