---
title: "Support"
author: "Kimberly A. Brink"
date: '`r format(Sys.Date(),"%B %d, %Y")`'
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

```{r install_libraries}
library(lubridate) #for calculating ages
library(Hmisc) #for reading SPSS files
library(pander) #for displaying tables
```

```{r standard_error}
#calculates standard error
s.error <- function(x) sd(x)/sqrt(length(x))
```

```{r barplot_medians_func}
#function that takes in a vector of median values, num of clusters, a vector of names for those clusters, a vector of names for groups, and a vector of colors for each group and formats them into a clustered bar graph
#if you do not have different groups, just set groupnames to NA
barplot_medians <- function(vals,nclusters,clusternames,groupnames,groupcolors){
  #formats medians for all test trials across all participants into a matrix
  medians <- matrix(vals,ncol=nclusters,byrow=TRUE)

  #label the columns with trial number
  colnames(medians) <- clusternames
  #label the rows with trial type
  rownames(medians) <- groupnames
  #convert matrix to table
  medians <- as.table(medians)
  
  #calculate maximum value for y-axis with added padding so that we can include value labels above the bars
  g.range = range(ifelse(min(medians)>0,0,floor(min(medians))-10),ceiling(max(medians))+10)
  
  #plot medians
  plot.new()
  barCenters <- barplot(medians, 
                        col=groupcolors, 
                        legend = ifelse(!is.na(groupnames),rownames(medians),FALSE),
                        ylab = "Median Looking Time (sec)", 
                        beside = TRUE, 
                        ylim=g.range)
  
  #print median values above bars
  text(barCenters,medians+1.5,label=medians)
}
```

```{r barplot_means_func}
#function that takes in a vector of mean values, s.error values, num of clusters, a vector of names for those clusters, a vector of names for groups, and a vector of colors for each group and formats them into a clustered bar graph with standard error bars
#if you do not have different groups, just set groupnames to NA
barplot_means <- function(meanvals,se.vals,nclusters,clusternames,groupnames,groupcolors){
  #formats means for all test trials across all participants into a matrix
  means <- matrix(meanvals,ncol=nclusters,byrow=TRUE)
  
  #formats standard error for all test trials across all participants into a matrix
  se <- matrix(se.vals,ncol=nclusters,byrow=TRUE)
  
  #label the columns with trial number
  colnames(means) <- clusternames
  #label the rows with trial type
  rownames(means) <- groupnames
  
  #convert matrix to table
  means <- as.table(means)
  #round mean values for printing above bars
  meansR <- round(means,2)
  
  #calculate maximum value for y-axis with added padding so that we can include value labels above the bars
  g.range = range(ifelse(min(means)>0,0,floor(min(means))-10),ceiling(max(means))+10)
  
  #plot means
  plot.new()
  barCenters <- barplot(means, 
                        col=groupcolors, 
                        legend.text = ifelse(!is.na(groupnames),rownames(means),FALSE),
                        ylab = "Mean Looking Time (sec)", 
                        beside = TRUE, 
                        ylim=g.range)
  
  #plot +/- 1 standard error bars
  arrows(barCenters, means-se, barCenters, means+se, lwd = 2, length=0.05, angle=90, code=3)
  
  #print mean values above bars
  text(barCenters,means+se+1.5,label=meansR)
}
```

```{r load_file}
filename="/Volumes/lsa-research01/ALL STUDIES/Current Studies/Support/Support 3.0/Support - Data v3.2.sav"
support <- spss.get(filename,datevars=c("DOB","DOT")) #recognize DOB and DOT files as date variables
```

```{r calculate_age}
#calculate age in months
support$Age = (year(as.period(interval(support$DOB, support$DOT)))*12) + month(as.period(interval(support$DOB, support$DOT))) + (day(as.period(interval(support$DOB, support$DOT)))/30)

#select only infants that participated after 2016-01-13 (the start date of the final design of the study)
support = support[which(support$DOT>='2016-01-13'),]
```

```{r calculate_crawl}
#converts variable strings to values for the crawl variable 
support$Crawl2 = ifelse(support$Crawl=="Very true", 3, 
                        ifelse(support$Crawl==levels(support$Crawl)[4], 2, #can't get text to match
                               ifelse(support$Crawl=="Rarely true", 1, 
                                      ifelse(support$Crawl=="Not at all true", 0, NA))))
```

```{r calculate_dishab}
#calculate dishabituation: change in looking time from last familiarization trial to first test trial
#Positive values mean that the infant renewed interest and dishabituated
#Negative values mean that the infant looked for less time during test and did not dishabituate

support$dishab = ifelse(support$FirstTest == "No Contact",
                        support$NoContactTest1-support$LeftFam2,
                        support$ContactTest1-support$RightFam2)

#calculate how much infant habituation during familiarization: decrease in looking time for first familiarization trial to last familiarization trial
#support$hab = ifelse(support$FirstTest == "No Contact",
#                       support$RightFam1-support$LeftFam2,
#                       support$LeftFam1-support$RightFam2)

#calculate whether infants habituated (looking time on last trial decreased to less than half looking time of first trial)
#support$habYN = ifelse(support$FirstTest == "No Contact",
#                        ifelse(support$RightFam1>2*support$LeftFam2,1,0),
#                        ifelse(support$LeftFam1>2*support$RightFam2,1,0))

#calculate whether infants habituated (looking time on last two trials decreased to less than half looking time of first two trial)
#support$habYN = ifelse(mean(support$RightFam1+support$LeftFam1)>mean(support$LeftFam2+support$RightFam2),1,0)
```

###Demographics.
There were `r length(support$SubID)` participants.

```{r computer_error}
#calculate the average looking time for each participant across all 3 'Contact' test trials
support$ContactTest.Average = (support$ContactTest1+support$ContactTest2+support$ContactTest3)/3

#calculate the average looking time for each participant across all 3 'No Contact' test trials
support$NoContactTest.Average = (support$NoContactTest1+support$NoContactTest2+support$NoContactTest3)/3

#Remove all participants that are missing data for any of the test trials
support = support[which(support$ContactTest.Average!='NA'),]
support = support[which(support$NoContactTest.Average!='NA'),]

#Remove all participants for which the data collection program unexpectedly closed during the study
support = support[which(support$Closed!="Closed"),]
```

Now there are `r length(support$SubID)` participants after removing for computer error. 

```{r crawling}
#support = support[which(support$Crawl2>0),]
```

Now there are `r length(support$SubID)` participants after removing for crawling. 

```{r demographics}
#divide participants into two data sets based on whether they were above or below the mean age
ageSplit = median(support$Age)
youngsupport = support[which(support$Age<=ageSplit),]
oldsupport = support[which(support$Age>ageSplit),]

#assign colors for plots for the two conditions (Contact and No Contact)
conCol= rgb(red=87/255,green=144/255,blue=154/255) #color for Contact condition
nocCol= rgb(red=237/255,green=186/255,blue=20/255) #color for No Contact condition

#print a table that summarizes age data
pander(summary(support$Age), caption="Age in months")
```

There are `r sum(support$Sex == "Female")` females in the sample.

The first date of test was `r min(support$DOT)`.

The most recent date of test was `r max(support$DOT)`.

#Tests for normality
```{r boxplots}
boxplot(support$ContactTest1, 
        support$ContactTest2, 
        support$ContactTest3, 
        support$NoContactTest1, 
        support$NoContactTest2, 
        support$ContactTest3, 
        col = c(rep(conCol,3),rep(nocCol,3)),
        ylab = 'Looking Time')
```

```{r histograms, fig.width = 8, fig.height = 5}
hist(support$ContactTest1, col = conCol, breaks = 10, main = "Contact Test 1", xlab="Looking Times")
hist(support$ContactTest2, col = conCol, breaks = 10, main = "Contact Test 2", xlab="Looking Times")
hist(support$ContactTest3, col = conCol, breaks = 10, main = "Contact Test 3", xlab="Looking Times")
hist(support$NoContactTest1, col = nocCol, breaks = 10, main = "No Contact Test 1", xlab="Looking Times")
hist(support$NoContactTest2, col = nocCol, breaks = 10, main = "No Contact Test 2", xlab="Looking Times")
hist(support$NoContactTest3, col = nocCol, breaks = 10, main = "No Contact Test 3", xlab="Looking Times")
```

#Non-Parametric Analyses
##Bar Graph: Median by Test Trial
```{r barplot_median_trial, fig.path='Figures/', dev=c('png','postscript'), fig.width = 6, fig.height = 6}
#formats medians for all test trials across all participants into a matrix
barplot_medians(vals=c(median(support$ContactTest1),
                       median(support$ContactTest2),
                       median(support$ContactTest3),
                       median(support$NoContactTest1),
                       median(support$NoContactTest2),
                       median(support$NoContactTest3)),
                nclusters=3,
                clusternames = c('Test Trial 1','Test Trial 2','Test Trial 3'),
                groupnames = c('Contact','No Contact'),
                groupcolors = c(conCol,nocCol))
```

```{r wilcox_test}
#perform nonparametric t-tests for each test trial and store values for printing in markdown
t1 = wilcox.test(support$NoContactTest1,support$ContactTest1,paired=TRUE)
#find z value
Z1 = round(qnorm(t1$p.value/2),digits = 2)
#find p-value
p1 = round(t1$p.value, digits = 2)
#convert p-value to APA style
p1 <- ifelse(p1<0.001, 0.001, ifelse(p1<0.01, 0.01, ifelse(p1<0.05,0.05,p1)))

t2 = wilcox.test(support$NoContactTest2,support$ContactTest2,paired=TRUE)
Z2 = round(qnorm(t2$p.value/2),digits = 2)
p2 = round(t2$p.value, digits = 2)
p2 <- ifelse(p2<0.001, 0.001, ifelse(p2<0.01, 0.01, ifelse(p2<0.05,0.05,p2)))

t3 = wilcox.test(support$NoContactTest3,support$ContactTest3,paired=TRUE)
Z3 = round(qnorm(t3$p.value/2),digits = 2)
p3 = round(t3$p.value, digits = 2)
p3 <- ifelse(p3<0.001, 0.001, ifelse(p3<0.01, 0.01, ifelse(p3<0.05,0.05,p3)))
```

There was no difference in looking times between the first Contact test event (Median =
`r median(support$ContactTest1)`) and the first No Contact event (Median = 
`r median(support$NoContactTest1)`),  (Wilcoxon *Z* = `r Z1`,  *p* = `r p1`)

There was no difference in looking times between the second Contact test event (Median = `r median(support$ContactTest2)`) and the second No Contact event (Median = 
`r median(support$NoContactTest2)`),  (Wilcoxon *Z* = `r Z2`,  *p* = `r p2`).

There was no difference in looking times between the third Contact test event (Median =
`r median(support$ContactTest3)`) and the third No Contact event (Median = 
`r median(support$NoContactTest3)`),  (Wilcoxon *Z* = `r Z3`,  *p* = `r p3`).

##Bar Graph: Median by Test Trial (Younger infants)
```{r barplot_median_young, fig.path='Figures/', dev=c('png','postscript'), fig.width = 6, fig.height = 6}
barplot_medians(vals=c(median(youngsupport$ContactTest1),
                       median(youngsupport$ContactTest2),
                       median(youngsupport$ContactTest3),
                       median(youngsupport$NoContactTest1),
                       median(youngsupport$NoContactTest2),
                       median(youngsupport$NoContactTest3)),
                nclusters=3,
                clusternames = c('Test Trial 1','Test Trial 2','Test Trial 3'),
                groupnames = c('Contact','No Contact'),
                groupcolors = c(conCol,nocCol))
```

```{r wilcox_test_young}
#perform nonparametric t-tests for each test trial and store values for printing in markdown

t1 = wilcox.test(youngsupport$NoContactTest1,youngsupport$ContactTest1,paired=TRUE)
Z1 = round(qnorm(t1$p.value/2),digits = 2)
p1 = round(t1$p.value, digits = 2)
p1 <- ifelse(p1<0.001, 0.001, ifelse(p1<0.01, 0.01, ifelse(p1<0.05,0.05,p1)))

t2 = wilcox.test(youngsupport$NoContactTest2,youngsupport$ContactTest2,paired=TRUE)
Z2 = round(qnorm(t2$p.value/2),digits = 2)
p2 = round(t2$p.value, digits = 2)
p2 <- ifelse(p2<0.001, 0.001, ifelse(p2<0.01, 0.01, ifelse(p2<0.05,0.05,p2)))

t3 = wilcox.test(youngsupport$NoContactTest3,youngsupport$ContactTest3,paired=TRUE)
Z3 = round(qnorm(t3$p.value/2),digits = 2)
p3 = round(t3$p.value, digits = 2)
p3 <- ifelse(p3<0.001, 0.001, ifelse(p3<0.01, 0.01, ifelse(p3<0.05,0.05,p3)))
```

For younger infants, there was no difference in looking times between the first Contact test event (Median =
`r median(youngsupport$ContactTest1)`) and the first No Contact event (Median = 
`r median(youngsupport$NoContactTest1)`),  (Wilcoxon *Z* = `r Z1`,  *p* = `r p1`)

For younger infants, there was no difference in looking times between the second Contact test event (Median = `r median(youngsupport$ContactTest2)`) and the second No Contact event (Median = 
`r median(youngsupport$NoContactTest2)`),  (Wilcoxon *Z* = `r Z2`,  *p* = `r p2`).

For younger infants, there was no difference in looking times between the third Contact test event (Median =
`r median(youngsupport$ContactTest3)`) and the third No Contact event (Median = 
`r median(youngsupport$NoContactTest3)`),  (Wilcoxon *Z* = `r Z3`,  *p* = `r p3`).

##Bar Graph: Median by Test Trial (Older infants)
```{r barplot_median_old, fig.path='Figures/', dev=c('png','postscript'), fig.width = 6, fig.height = 6}
barplot_medians(vals=c(median(oldsupport$ContactTest1),
                       median(oldsupport$ContactTest2),
                       median(oldsupport$ContactTest3),
                       median(oldsupport$NoContactTest1),
                       median(oldsupport$NoContactTest2),
                       median(oldsupport$NoContactTest3)),
                nclusters=3,
                clusternames = c('Test Trial 1','Test Trial 2','Test Trial 3'),
                groupnames = c('Contact','No Contact'),
                groupcolors = c(conCol,nocCol))
```

```{r wilcox_test_old}
#perform nonparametric t-tests for each test trial and store values for printing in markdown

t1 = wilcox.test(oldsupport$NoContactTest1,oldsupport$ContactTest1,paired=TRUE)
Z1 = round(qnorm(t1$p.value/2),digits = 2)
p1 = round(t1$p.value, digits = 2)
p1 <- ifelse(p1<0.001, 0.001, ifelse(p1<0.01, 0.01, ifelse(p1<0.05,0.05,p1)))

t2 = wilcox.test(oldsupport$NoContactTest2,oldsupport$ContactTest2,paired=TRUE)
Z2 = round(qnorm(t2$p.value/2),digits = 2)
p2 = round(t2$p.value, digits = 2)
p2 <- ifelse(p2<0.001, 0.001, ifelse(p2<0.01, 0.01, ifelse(p2<0.05,0.05,p2)))

t3 = wilcox.test(oldsupport$NoContactTest3,oldsupport$ContactTest3,paired=TRUE)
Z3 = round(qnorm(t3$p.value/2),digits = 2)
p3 = round(t3$p.value, digits = 2)
p3 <- ifelse(p3<0.001, 0.001, ifelse(p3<0.01, 0.01, ifelse(p3<0.05,0.05,p3)))
```

There was no difference in looking times between the first Contact test event (Median =
`r median(oldsupport$ContactTest1)`) and the first No Contact event (Median = 
`r median(oldsupport$NoContactTest1)`),  (Wilcoxon *Z* = `r Z1`,  *p* = `r p1`)

There was no difference in looking times between the second Contact test event (Median = `r median(oldsupport$ContactTest2)`) and the second No Contact event (Median = 
`r median(oldsupport$NoContactTest2)`),  (Wilcoxon *Z* = `r Z2`,  *p* = `r p2`).

There was no difference in looking times between the third Contact test event (Median =
`r median(oldsupport$ContactTest3)`) and the third No Contact event (Median = 
`r median(oldsupport$NoContactTest3)`),  (Wilcoxon *Z* = `r Z3`,  *p* = `r p3`).

#Parametric Analyses
##Bar Graph: Means by Test Trial (All infants)
```{r barplot_means_trial, fig.path='Figures/', dev=c('png','postscript'), fig.width = 8, fig.height = 7}
barplot_means(meanvals = c(mean(support$ContactTest1),
                           mean(support$ContactTest2),
                           mean(support$ContactTest3),
                           mean(support$NoContactTest1),
                           mean(support$NoContactTest2),
                           mean(support$NoContactTest3)),
              se.vals = c(s.error(support$ContactTest1),
                          s.error(support$ContactTest2),
                          s.error(support$ContactTest3),
                          s.error(support$NoContactTest1),
                          s.error(support$NoContactTest2),
                          s.error(support$NoContactTest3)),
              nclusters = 3,
              clusternames = c("Test Trial 1", "Test Trial 2","Test Trial 3"),
              groupnames = c("Contact","No Contact"),
              groupcolors = c(conCol,nocCol))
```

```{r ttest}
t1 = t.test(support$NoContactTest1,support$ContactTest1,paired=TRUE)

t2 = t.test(support$NoContactTest2,support$ContactTest2,paired=TRUE)

t3 = t.test(support$NoContactTest3,support$ContactTest3,paired=TRUE)
```

There was no difference in looking times between the first Possible test event (M =
`r round(mean(support$ContactTest1), 2)`) and the first Impossible event (M = 
`r round(mean(support$NoContactTest1),2)`),  (t(`r round(t1$parameter,2)`) = `r round(t1$statistic,2)`,  *p* = `r round(t1$p.value,2)`).

There was no difference in looking times between the second Possible test event (M = 
`r round(mean(support$ContactTest2),2)`) and the second Impossible event (M = 
`r round(mean(support$NoContactTest2),2)`),  (t(`r round(t2$parameter,2)`) = `r round(t2$statistic,2)`,  *p* = `r round(t2$p.value,2)`).

There was no difference in looking times between the third Possible test event (M =
`r round(mean(support$ContactTest3),2)`) and the third Impossible event (M = 
`r round(mean(support$NoContactTest3),2)`),  (t(`r round(t3$parameter,2)`) = `r round(t3$statistic,2)`,  *p* = `r round(t3$p.value,2)`).

#Bar Graph: Means by Test Trial (Younger infants)
```{r young_means,fig.path='Figures/', dev=c('png','postscript'), fig.width = 6, fig.height = 6}
barplot_means(meanvals = c(mean(youngsupport$ContactTest1),
                           mean(youngsupport$ContactTest2),
                           mean(youngsupport$ContactTest3),
                           mean(youngsupport$NoContactTest1),
                           mean(youngsupport$NoContactTest2),
                           mean(youngsupport$NoContactTest3)),
              se.vals = c(s.error(youngsupport$ContactTest1),
                          s.error(youngsupport$ContactTest2),
                          s.error(youngsupport$ContactTest3),
                          s.error(youngsupport$NoContactTest1),
                          s.error(youngsupport$NoContactTest2),
                          s.error(youngsupport$NoContactTest3)),
              nclusters = 3,
              clusternames = c("Test Trial 1", "Test Trial 2","Test Trial 3"),
              groupnames = c("Contact","No Contact"),
              groupcolors = c(conCol,nocCol))
```

```{r ttest_young}
t1 = t.test(youngsupport$NoContactTest1,youngsupport$ContactTest1,paired=TRUE)

t2 = t.test(youngsupport$NoContactTest2,youngsupport$ContactTest2,paired=TRUE)

t3 = t.test(youngsupport$NoContactTest3,youngsupport$ContactTest3,paired=TRUE)
```

For infants younger than `r round(ageSplit,2)` months, there was no difference in looking times between the first Possible test event (M =
`r round(mean(youngsupport$ContactTest1), 2)`) and the first Impossible event (M = 
`r round(mean(youngsupport$NoContactTest1),2)`),  (t(`r round(t1$parameter,2)`) = `r round(t1$statistic,2)`,  *p* = `r round(t1$p.value,2)`).

For infants younger than `r round(ageSplit,2)` months, there was no difference in looking times between the second Possible test event (M = 
`r round(mean(youngsupport$ContactTest2),2)`) and the second Impossible event (M = 
`r round(mean(youngsupport$NoContactTest2),2)`),  (t(`r round(t2$parameter,2)`) = `r round(t2$statistic,2)`,  *p* = `r round(t2$p.value,2)`).

For infants younger than `r round(ageSplit,2)` months, there was no difference in looking times between the third Possible test event (M =
`r round(mean(youngsupport$ContactTest3),2)`) and the third Impossible event (M = 
`r round(mean(youngsupport$NoContactTest3),2)`),  (t(`r round(t3$parameter,2)`) = `r round(t3$statistic,2)`,  *p* = `r round(t3$p.value,2)`).

#Bar Graph: Means by Test Trial (Older infants)
```{r old_means,fig.path='Figures/', dev=c('png','postscript'), fig.width = 6, fig.height = 6}
barplot_means(meanvals = c(mean(oldsupport$ContactTest1),
                           mean(oldsupport$ContactTest2),
                           mean(oldsupport$ContactTest3),
                           mean(oldsupport$NoContactTest1),
                           mean(oldsupport$NoContactTest2),
                           mean(oldsupport$NoContactTest3)),
              se.vals = c(s.error(oldsupport$ContactTest1),
                          s.error(oldsupport$ContactTest2),
                          s.error(oldsupport$ContactTest3),
                          s.error(oldsupport$NoContactTest1),
                          s.error(oldsupport$NoContactTest2),
                          s.error(oldsupport$NoContactTest3)),
              nclusters = 3,
              clusternames = c("Test Trial 1", "Test Trial 2","Test Trial 3"),
              groupnames = c("Contact","No Contact"),
              groupcolors = c(conCol,nocCol))
```

```{r ttest_old}
t1 = t.test(oldsupport$NoContactTest1,oldsupport$ContactTest1,paired=TRUE)

t2 = t.test(oldsupport$NoContactTest2,oldsupport$ContactTest2,paired=TRUE)

t3 = t.test(oldsupport$NoContactTest3,oldsupport$ContactTest3,paired=TRUE)
```

For infants older than `r round(ageSplit,2)` months, there was no difference in looking times between the first Possible test event (M =
`r round(mean(oldsupport$ContactTest1), 2)`) and the first Impossible event (M = 
`r round(mean(oldsupport$NoContactTest1),2)`),  (t(`r round(t1$parameter,2)`) = `r round(t1$statistic,2)`,  *p* = `r round(t1$p.value,2)`).

For infants older than `r round(ageSplit,2)` months, there was no difference in looking times between the second Possible test event (M = 
`r round(mean(oldsupport$ContactTest2),2)`) and the second Impossible event (M = 
`r round(mean(oldsupport$NoContactTest2),2)`),  (t(`r round(t2$parameter,2)`) = `r round(t2$statistic,2)`,  *p* = `r round(t2$p.value,2)`).

For infants older than `r round(ageSplit,2)` months, there was no difference in looking times between the third Possible test event (M =
`r round(mean(oldsupport$ContactTest3),2)`) and the third Impossible event (M = 
`r round(mean(oldsupport$NoContactTest3),2)`),  (t(`r round(t3$parameter,2)`) = `r round(t3$statistic,2)`,  *p* = `r round(t3$p.value,2)`).


##Bar Graph: Average Looking Time Across Condition
```{r barplot_means, fig.path='Figures/', dev=c('png','postscript'), fig.width = 4, fig.height = 7, include=TRUE}
barplot_means(meanvals = c(mean(support$ContactTest.Average), 
                           mean(support$NoContactTest.Average)),
              se.vals = c(s.error(support$ContactTest.Average),
                          s.error(support$NoContactTest.Average)),
              nclusters = 2,
              clusternames = c("Contact","No Contact"),
              groupnames = NA,
              groupcolors = c(conCol,nocCol))

t.avg = t.test(support$ContactTest.Average,support$NoContactTest.Average,paired=T)
```

There was no difference in looking times between the average looking time of all three Possible test events (M =
`r round(mean(support$ContactTest.Average),2)`) and the average looking time of all three Impossible events (M = 
`r round(mean(support$NoContactTest.Average),2)`),  (t(`r round(t.avg$parameter,2)`) = `r round(t.avg$statistic,2)`,  *p* = `r round(t.avg$p.value,2)`).

#Between-Subject Analyses
##Bar Graph: Medians of first trial (Between-Subject)
```{r between_median}
imp.first <- support[which(support$FirstTest == "No Contact"),]
poss.first <- support[which(support$FirstTest == "Contact"),]


barplot_medians(vals = c(median(poss.first$ContactTest1), 
                         median(imp.first$NoContactTest1)),
                nclusters = 2, 
                clusternames = c("Contact","No Contact"),
                groupnames = NA,
                groupcolors = c(conCol,nocCol))

t1 = wilcox.test(imp.first$NoContactTest1,poss.first$ContactTest1,paired=FALSE)
Z1 = round(qnorm(t1$p.value/2),digits = 2)
p1 = round(t1$p.value, digits = 2)
p1 <- ifelse(p1<0.001, 0.001, ifelse(p1<0.01, 0.01, ifelse(p1<0.05,0.05,p1)))
```

There was no difference in looking times between the first Contact test event (Median =
`r median(poss.first$ContactTest1)`) and the first No Contact event (Median = 
`r median(imp.first$NoContactTest1)`) between subjects,  (Wilcoxon *Z* = `r Z1`,  *p* = `r p1`)

##Bar Graph: Means of first trial (Between-Subject)
```{r between_mean}
barplot_means(meanvals = c(mean(poss.first$ContactTest1), 
                           mean(imp.first$NoContactTest1)),
              se.vals = c(s.error(poss.first$ContactTest1),
                          s.error(imp.first$NoContactTest1)),
              nclusters = 2,
              clusternames = c("Contact","No Contact"),
              groupnames = NA,
              groupcolors = c(conCol,nocCol))

t1 = t.test(imp.first$NoContactTest1,poss.first$ContactTest1,paired=FALSE)
```

There was no difference in looking times between the first Possible test event (M =
`r round(mean(poss.first$ContactTest1), 2)`) and the first Impossible event (M = 
`r round(mean(imp.first$NoContactTest1),2)`) between subjects,  (t(`r round(t1$parameter,2)`) = `r round(t1$statistic,2)`,  *p* = `r round(t1$p.value,2)`).

#Dishabituation Analyses
###(Difference between last familiarization and first test)
Positive values indicate dishabituation (renewed interest), negative values indicate no dishabituation

```{r dishab_hist}
hist(poss.first$dishab, col = conCol, breaks = 10, main = "Dishabituation to first Possible Test Trial", xlab = "Dishabituation")
hist(imp.first$dishab, col = nocCol, breaks = 10, main = "Dishabituation to first Impossible Test Trial", xlab = "Dishabituation")
```

##Bar Graph: Median of Dishabituation
```{r dishab_median}
barplot_medians(vals = c(median(poss.first$dishab), median(imp.first$dishab)),
                nclusters = 2,
                clusternames = c("Contact","No Contact"),
                groupnames = NA,
                groupcolors = c(conCol,nocCol))

t1 = wilcox.test(imp.first$dishab,poss.first$dishab)
Z1 = round(qnorm(t1$p.value/2),digits = 2)
p1 = round(t1$p.value, digits = 2)
p1 <- ifelse(p1<0.001, 0.001, ifelse(p1<0.01, 0.01, ifelse(p1<0.05,0.05,p1)))
```

There was no difference in dishabituation to the first Contact test event (Median =
`r median(poss.first$dishab)`) and dishabituation to the first No Contact event (Median = 
`r median(imp.first$dishab)`),  (Wilcoxon *Z* = `r Z1`,  *p* = `r p1`)

##Bar Graph: Mean of dishabituation
```{r dishab_mean}
barplot_means(meanvals = c(mean(poss.first$dishab), 
                           mean(imp.first$dishab)),
              se.vals = c(s.error(poss.first$dishab),
                          s.error(imp.first$dishab)),
              nclusters = 2,
              clusternames = c("Contact","No Contact"),
              groupnames = NA,
              groupcolors = c(conCol,nocCol))

t1 = t.test(imp.first$dishab,poss.first$dishab)
```

There was no difference in dishabituation to the first Possible test event (M =
`r round(mean(poss.first$dishab), 2)`) and dishabituation to the first Impossible event (M = 
`r round(mean(imp.first$dishab),2)`),  (t(`r round(t1$parameter,2)`) = `r round(t1$statistic,2)`,  *p* = `r round(t1$p.value,2)`).

##Breakdown of young (<=`r round(ageSplit,2)` months) vs old infants (> `r round(ageSplit,2)`)
```{r young_vs_old_data}
pander(table(support$Age<=ageSplit, support$Order))
pander(table(support$Age<=ageSplit, support$Sex))

pander(table(youngsupport$Sex,youngsupport$Order), caption=paste("Infants younger than", round(ageSplit,2), "months"))
pander(table(oldsupport$Sex,oldsupport$Order), caption=paste("Infants older than", round(ageSplit,2), "months"))
```

There were `r length(youngsupport$ContactTest.Average)` less than or equal to `r round(ageSplit,2)` months old.
There were `r length(oldsupport$ContactTest.Average)` greater than or equal to `r round(ageSplit,2)` months old.

#Bar Graph: Mean Looking Time across Trials (Young vs Old)
```{r young_vs_old, fig.path='Figures/', dev=c('png','postscript'), fig.width = 6.5, fig.height = 6}
barplot_means(meanvals = c(mean(youngsupport$ContactTest.Average), 
                           mean(oldsupport$ContactTest.Average),
                           mean(youngsupport$NoContactTest.Average),
                           mean(oldsupport$NoContactTest.Average)),
              se.vals = c(s.error(youngsupport$ContactTest.Average), 
                          s.error(oldsupport$ContactTest.Average), 
                          s.error(youngsupport$NoContactTest.Average),
                          s.error(oldsupport$NoContactTest.Average)),
              nclusters = 2,
              clusternames = c(paste("<",round(ageSplit,2),"months"),
                               paste(">",round(ageSplit,2),"months")),              
              groupnames = c("Contact", "No Contact"),
              groupcolors = c(conCol,nocCol))

t.avg.old = t.test(oldsupport$ContactTest.Average,oldsupport$NoContactTest.Average,paired=T)
t.avg.young = t.test(youngsupport$ContactTest.Average,youngsupport$NoContactTest.Average,paired=T)
```


For infants younger than `r round(ageSplit,2)` months, there was no difference in looking times between the second Possible test event (M = 
`r round(mean(youngsupport$ContactTest.Average),2)`) and the second Impossible event (M = 
`r round(mean(youngsupport$NoContactTest.Average),2)`),  (t(`r round(t.avg.young$parameter,2)`) = `r round(t.avg.young$statistic,2)`,  *p* = `r round(t.avg.young$p.value,2)`).

For infants older than `r round(ageSplit,2)` months, there was no difference in looking times between the first Possible test event (M =
`r round(mean(oldsupport$ContactTest.Average), 2)`) and the first Impossible event (M = 
`r round(mean(oldsupport$NoContactTest.Average),2)`),  (t(`r round(t.avg.old$parameter,2)`) = `r round(t.avg.old$statistic,2)`,  *p* = `r round(t.avg.old$p.value,2)`).


#Bar Graph: Mean Looking Time on First Trial (Young vs Old)
```{r young_vs_old_t1,fig.path='Figures/', dev=c('png','postscript'), fig.width = 6.5, fig.height = 6}
barplot_means(meanvals = c(mean(youngsupport$ContactTest1), 
                           mean(oldsupport$ContactTest1), 
                           mean(youngsupport$NoContactTest1),
                           mean(oldsupport$NoContactTest1)),
              se.vals = c(s.error(youngsupport$ContactTest1), 
                          s.error(oldsupport$ContactTest1), 
                          s.error(youngsupport$NoContactTest1), 
                          s.error(oldsupport$NoContactTest1)),
              nclusters = 2,
              clusternames = c(paste("<",round(ageSplit,2),"months"),
                               paste(">",round(ageSplit,2),"months")),              
              groupnames = c("Contact", "No Contact"),
              groupcolors = c(conCol,nocCol))

t1.young = t.test(youngsupport$NoContactTest1,youngsupport$ContactTest1,paired=TRUE)
t1.old = t.test(oldsupport$NoContactTest1,oldsupport$ContactTest1,paired=TRUE)
```

For infants younger than `r round(ageSplit,2)` months, there was no difference in looking times between the first Possible test event (M =
`r round(mean(youngsupport$ContactTest1), 2)`) and the first Impossible event (M = 
`r round(mean(youngsupport$NoContactTest1),2)`),  (t(`r round(t1.young$parameter,2)`) = `r round(t1.young$statistic,2)`,  *p* = `r round(t1.young$p.value,2)`).

For infants older than `r round(ageSplit,2)` months, there was no difference in looking times between the first Possible test event (M =
`r round(mean(oldsupport$ContactTest1), 2)`) and the first Impossible event (M = 
`r round(mean(oldsupport$NoContactTest1),2)`),  (t(`r round(t1.old$parameter,2)`) = `r round(t1.old$statistic,2)`,  *p* = `r round(t1.old$p.value,2)`).

#Bar Graph: Mean Looking Time on Third Trial (Young vs Old)
```{r young_vs_old_t3,fig.path='Figures/', dev=c('png','postscript'), fig.width = 6.5, fig.height = 6}
barplot_means(meanvals = c(mean(youngsupport$ContactTest3), 
                           mean(oldsupport$ContactTest3),
                           mean(youngsupport$NoContactTest3),
                           mean(oldsupport$NoContactTest3)),
              se.vals = c(s.error(youngsupport$ContactTest3), 
                          s.error(oldsupport$ContactTest3),
                          s.error(youngsupport$NoContactTest3),
                          s.error(oldsupport$NoContactTest3)),
              nclusters = 2,
              clusternames = c(paste("<",round(ageSplit,2),"months"),
                               paste(">",round(ageSplit,2),"months")),              
              groupnames = c("Contact", "No Contact"),
              groupcolors = c(conCol,nocCol))

t3.young = t.test(youngsupport$NoContactTest3,youngsupport$ContactTest3,paired=TRUE)
t3.old = t.test(oldsupport$NoContactTest3,oldsupport$ContactTest3,paired=TRUE)

```

For infants younger than `r round(ageSplit,2)` months, there was no difference in looking times between the third Possible test event (M =
`r round(mean(youngsupport$ContactTest3), 2)`) and the first Impossible event (M = 
`r round(mean(youngsupport$NoContactTest3),2)`),  (t(`r round(t3.young$parameter,2)`) = `r round(t3.young$statistic,2)`,  *p* = `r round(t3.young$p.value,2)`).

For infants older than `r round(ageSplit,2)` months, there was no difference in looking times between the third Possible test event (M =
`r round(mean(oldsupport$ContactTest3), 2)`) and the first Impossible event (M = 
`r round(mean(oldsupport$NoContactTest3),2)`),  (t(`r round(t3.old$parameter,2)`) = `r round(t3.old$statistic,2)`,  *p* = `r round(t3.old$p.value,2)`).
